Solving an Exam Scheduling Problem Using a Genetic Algorithm

### 2009-12-02 trunk/Paper/paper.tex
Updated paper with some new graphs
Paper update

\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{url}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDFMARK for TeX and GhostScript
% Uncomment and complete the following for metadata if
% your paper is typeset using TeX and GhostScript (e.g
% if you use .ps or .eps files in your paper):
% \special{! /pdfmark where
% {pop} {userdict /pdfmark /cleartomark load put} ifelse
% [ /Author (John Doe, Jane Doe)
% /Title (Paper Title)
% /Keywords (AAAI, artificial intelligence)
% /DOCINFO pdfmark}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDFINFO for PDFTeX
% Uncomment and complete the following for metadata if
% your paper is typeset using PDFTeX
\pdfinfo{
  /Title (Solving an Exam Scheduling Problem Using a Genetic Algorithm)
  /Subject (Input The Proceedings Title Here)
  /Author (Dave, Kordalewski;
           Caigu, Liu;
           Kevin, Salvesen;)
} 
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Uncomment only if you need to use section numbers
% and change the 0 to a 1 or 2
% \setcounter{secnumdepth}{0}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

  A schedule, then, is a mapping from courses to rooms at times, 
  which we can express like this:

  \[ C \rightarrow \left\{(r,\ a,\ b)\ |\ r\in R,\ a\in \{1..d\},\ b\in \{1..t\}\right\} \]

  The total number of different possible schedules in any problem instance is
  $|C|^{|R|dt}$ which is typically far too large for brute force search.
  For example, one instance that we examine (scheduling the fall semester
  exams at the University of Toronto in 2009) involves 603 courses, 7 days,
  8 times, 43 rooms, and 21945 students. This instance admits approximately 
  $10^{6695}$ different possible schedules.

  Typically, with relatively loose constraints on the number of rooms and 
  particular schedules of students, it is not difficult to find some consistent
  schedule; that is, one where no student is asked to write 2 exams
  simultaneously and no room has 2 exams occur in it simultaneously.

@@

  The timetable that a student $s \in S$ has under some particular schedule 
  $K$ may be represented in this way:

  \[ TT_s(K) = \left\{(a, b)\ |\ \exists r \in R, \exists c \in L_s, K(c)=(r, a, b) \right\} \]

  which may, in general, be a multiset.

  Similarly, the timetable for a room $r \in R$ is

  \[ TT_r(K)=\left\{(a, b)\ |\ \exists c\in C,\ K(c)=(r, a, b)\right\} \]

  We define two quality functions, mapping schedules to real numbers in [0,1],
  one for students and one for rooms. These are meant to capture how much they
  "like" the schedule under consideration. (For instance, a student will rate
  poorly any schedule where she has 2 consecutive exams, and very poorly any 
  schedule which expects her to take two exams at the same time.)

  \[ Q_s(K)=q_{student}\left(TT_s(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
  \[ Q_r(K)=q_{room}\left(TT_r(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]

  We will consider later how to define these functions in a reasonable way.

  The quality of a schedule, then, is given by

  \[ Q(K)= \frac{\left(w_s\left(\sum_{s \in S}{\frac{Q_s(K)}{|S|}}\right) +
     w_r\left(\sum_{r\in R}{\frac{Q_r(K)}{|R|}} \right)\right)}{(w_s+w_r)} \]

  where $w_s$ and $w_r$ are the weightings we can use to indicate that we 
  care somewhat more about students preferences than rooms, giving a quality
  (or fitness, in the terminology of genetic algorithms) in the range [0,1] for
  any schedule.

@@

\section{Approach and Implementation}
  \subsection{Differences From Natural Exam Scheduling Problems}
    The problem we have constructed here is a compromise between the most
    mathematically pure problem with the same sort of properties (which might ignore the
    rooms and related constraints completely) and the problem that must be solved in the
    real world when scheduling thousands of students, which has some important differences
    from this problem.

  \subsection{Tools used}
    The program is implemented in Java, to generate possible schedules and to calculate
    their fitness. A Mathematica program is also written to plot graphs, given the data
    generated by the Java program. Additionally, we use Google code 
    subversion\footnote{\url{http://code.google.com/p/csc384-genetic-algorithms-project/}}
    and Google doc to share the work.

  \subsection{Fitness Function}
    In order to assign each schedule a fitness, we need to define the functions 
    $q_{student}$ and $q_{room}$. They need to capture the idea of determining how
    much each student or room likes their given timetable. We do this by calculating 
    the penalty incurred by a room, which is a measure of it having properties that 
    are unpleasant to the student or room, and then using the formula

    \[ q(TT)=\frac{1}{1+penalty_{TT}} \]

    This function will be in the desired range or [0,1].

    For students, we accrue a penalty of 50 for every time that is repeated in a 
    timetable, 5 when there are exams in 2 consecutive times, 3 when 2 exams happen 
    on the same day, and 0.5 when the student has exams on consecutive days. This 
    accords reasonably well with real students' hopes for their exam timetables.

    Rooms have the same penalty of 50 when it is used by two exams at the same time, 
    5 if it hosts exams in consecutive time slots (reasoning that the invigilators 
    need time to let the students out and in and prepare for the next exam) and a 
    penalty of 0.5 when a room is empty for an entire day between uses (reasoning 
    that the room requires some amount of modifications to be used for exams, and 
    it is difficult to use it for other purposes if it will still be used for exams 
    soon.)

    The details of the function we used are essentially arbitrary.

  \subsection{Genetic Algorithm Search}
    Although there are many readily available open source genetic algorithm software
    packages available (see JGAP \cite{website:JGAP}, Jenes\cite{website:jenes}) we 
    decided to write a simple, yet general, genetic algorithm package, in Java, that
    can interface with data types designed by others, so long as they implement a few
    necessary methods. The package we wrote is more than adequate for investigating
    the scheduling problem under consideration here.

    \subsubsection{General Description}
      The genetic algorithm maintains a population of schedules and constructs subsequent
      generations by adding new random schedules, making copies of high quality schedules 
      in the present generation, applies the mutation operator to schedules, or uses the
      crossover operator on 2 schedules. When performing mutation or crossover, the 
      schedules are selected with a probability proportional to their fitness, allowing 
      the schedules with high quality tobe selected more frequently and make more 
      "offspring".

      The free parameters in our GA are the size of the population, number of
      generations to simulate, and proportions of new generations toi be made by the copy, 
      random, mutation, and crossover methods.

    \subsubsection{Representation}
      We use the naive representation of schedules. A schedule contains lists of size $|C|$
      containing the room in which each course's exam is to occur and the timing
      (corresponding to a pair or integers representing the day and time) of each course's
      exam.

@@

      The neighbouring schedules are those where only one course meets in either a different
      room, on a different day, or at a different time. This leads to a total of
      $ |C|(d+t+|R|-3) $ schedules in the neighbourhood of any given schedule.

      It can be seen that this method is completely deterministic given a starting schedule,
      and will always return a local maximum

    \subsubsection{Mutate Search}
      Mutate Search starts by generating a random schedule. It makes an alternative schedule
      (by using the same mutate operator made for the genetic algorithm) with 1\% of the
      data randomized and chooses which of the schedules is better. This operation is
      repeated until it has performed as many evaluations as it is allowed to.

@@

  \subsection{Options for the Genetic Algorithm}
    \subsubsection{Scheduling Problem Instances}
      We generated 2 sets of 3 scheduling problem instances to test the algorithms above with. 
      The package ScheduleInstanceGenerator in our code repository manages the generation of 
      random problem instances with many free parameters to specify characteristics of the 
      output instance.

      The following table characterized the problem instances that we used:

      \begin{tabular}{|c|c|c|c|c|c|c|}
      \hline  & Tiny & Medium & Large & Tiny /w Cor. & Med /w Cor. & Large /w Cor. \\ 
      \hline d & 7 & 10 & 7 & 7 & 10 & 7 \\ 
      \hline t & 5 & 8 & 8 & 5 & 8 & 8 \\ 
      \hline $|C|$ & 20 & 200 & 603 & 20 & 200 & 600 \\ 
      \hline $|R|$ & 4 & 10 & 43 & 4 & 10 & 43 \\ 
      \hline $|S|$ & 50 & 300 & 21945 & 50 & 300 & 21945 \\ 
      \hline $|L_s|$ & 1-5 & 1-6 & 1-5 & 2-6 & 2-6 & 2-6 \\ 
      \hline \# Course Types & N/A & N/A & N/A & 4 & 10 & 20 \\ 
      \hline 
	\end{tabular}

	  The full input files for these instances are available in our repository in the 
	  directory "Instance Input Files".

    \subsubsection{Selecting Parameters for the GA}
      [\ldots]
  \subsection{Comparison Between the Different Searches}
    [\ldots]

\section{Evaluation}
  \subsection{Success}
    [\ldots]
    Seeing as our fitness function gives a very high penalty for schedules that have direct
    conflicts (either a student with two exams at the same moment, or two different exams
    in a same room at the same moment), if such a schedule exists given the input data, our
    algorithms will very quickly tend to reject all such schedules.

\section{Conclusion}
  We come up with a conclusion that in this particular problem, genetic algorithm
  is better than random search, and it is better than hill-climbing if the input 
  is a large data set. But it may not be the best solution for this problem, since
  the Mutate Search performs way better.

\bibliographystyle{aaai}
\bibliography{paperBib}
\end{document}

>>>>

\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{url}
\usepackage[table]{xcolor}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{float}
\usepackage{natbib}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDFMARK for TeX and GhostScript
% Uncomment and complete the following for metadata if
% your paper is typeset using TeX and GhostScript (e.g
% if you use .ps or .eps files in your paper):
% \special{! /pdfmark where
% {pop} {userdict /pdfmark /cleartomark load put} ifelse
% [ /Author (John Doe, Jane Doe)
% /Title (Paper Title)
% /Keywords (AAAI, artificial intelligence)
% /DOCINFO pdfmark}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PDFINFO for PDFTeX
% Uncomment and complete the following for metadata if
% your paper is typeset using PDFTeX
\pdfinfo{
  /Title (Solving an Exam Scheduling Problem Using a Genetic Algorithm)
  /Subject (Input The Proceedings Title Here)
  /Author (Dave, Kordalewski;
           Caigu, Liu;
           Kevin, Salvesen;)
} 
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Uncomment only if you need to use section numbers
% and change the 0 to a 1 or 2
% \setcounter{secnumdepth}{0}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

  A schedule, then, is a mapping from courses to rooms at times, 
  which we can express like this:
  \[ C \rightarrow \left\{(r,\ a,\ b)\ |\ r\in R,\ a\in \{1..d\},\ b\in \{1..t\}\right\} \]
  Typically, with relatively loose constraints on the number of rooms and 
  particular schedules of students, it is not difficult to find some consistent
  schedule; that is, one where no student is asked to write 2 exams
  simultaneously and no room has 2 exams occur in it simultaneously.

@@

  The timetable that a student $s \in S$ has under some particular schedule 
  $K$ may be represented in this way:
  \[ TT_s(K) = \left\{(a, b)\ |\ \exists r \in R, \exists c \in L_s, K(c)=(r, a, b) \right\} \]
  which may, in general, be a multiset.

  Similarly, the timetable for a room $r \in R$ is
  \[ TT_r(K)=\left\{(a, b)\ |\ \exists c\in C,\ K(c)=(r, a, b)\right\} \]

  We define two quality functions, mapping schedules to real numbers in [0,1],
  one for students and one for rooms. These are meant to capture how much they
  ``like" the schedule under consideration. (For instance, a student will rate
  poorly any schedule where she has 2 consecutive exams, and very poorly any 
  schedule which expects her to take two exams at the same time. A room will 
  rate poorly a schedule that assigns two different exams to be held in it at 
  the same moment).
  \[ Q_s(K)=q_{student}\left(TT_s(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
  \[ Q_r(K)=q_{room}\left(TT_r(K)\right):\textit{ Schedules}\rightarrow [0,\ 1] \]
  We will consider later how to define these functions in a reasonable way.

  The quality of a schedule, then, is given by
  \[ Q(K)= \frac{\left(w_s\left(\sum_{s \in S}{\frac{Q_s(K)}{|S|}}\right) +
     w_r\left(\sum_{r\in R}{\frac{Q_r(K)}{|R|}} \right)\right)}{(w_s+w_r)} \]
  where $w_s$ and $w_r$ are the weightings we can use to indicate that we 
  care somewhat more about students preferences than rooms, giving a quality
  (or fitness, in the terminology of genetic algorithms) in the range [0,1] for
  any schedule.

@@

\section{Approach and Implementation}
  \subsection{Differences From Natural Exam Scheduling}
    The problem we have constructed here is a compromise between more mathematically
    pure problems with the same sort of properties (which might ignore the rooms' 
    constraints completely) and the problem that must be solved in the real world 
    when scheduling thousands of students, which has some important differences from 
    this problem. For example, we don't deal with constraints like room capacity or 
    exams with varying length.

  \subsection{Tools used}
    Our application was entirely written in Java. We also wrote another Java application 
    to generate problem instance files for testing our algorithms. We used Mathematica 
    to plot all of our graphs.  Additionally, we use Google code's subversion
    repository\footnote{\url{http://code.google.com/p/csc384-genetic-algorithms-project/}},
    Google docs and Google Wave to collaborate on this report.

  \subsection{Fitness Function}
    In order to assign each schedule a fitness, we need to define the functions
    $q_{student}$ and $q_{room}$. They need to capture the idea of determining 
    how much each student or room likes their given timetable. We do this by giving 
    penalties to schedules when they have properties that are disliked by the 
    student or room, and then using the formula
    \[ q(TT)=\frac{1}{1+penalty_{TT}} \]
    This function will be in the desired range of [0,1].

    We assign 50 point penalties for inconsistent schedules, 3-5 points when 
    students are assigned multiple exams on the same day, and 0.5 when students 
    don't have a rest day between exams. Rooms have the same stiff 50 point penalty 
    for inconsistencies, 5 for hosting exams in consecutive time slots, and 0.5 for 
    any days without exams between the days it is first used and last used.

    We chose the weighting constants $w_s=2$ and $w_r=1$, which suggest that we are 
    more concerned with the students' schedules than the rooms'.

    The details of the function we used are essentially arbitrary, and we expect 
    similar results would be attained for any other penalty-based fitness function 
    modelling the preferences of students and requirements of rooms.

  \subsection{Genetic Algorithm Search}
    Although there are many readily available open source genetic algorithm software 
    packages available (see JGAP, Jenes) we decided to write a simple, yet general, 
    genetic algorithm package, in Java, that can interface with data types designed 
    by others, so long as they implement a few necessary methods. The package we 
    wrote is more than adequate for investigating the scheduling problem under 
    consideration here.

    \subsubsection{General Description}
      The genetic algorithm maintains a population of schedules and constructs subsequent 
      generations by adding new random schedules, making copies of high quality schedules 
      in the present generation, applies the mutation operator to schedules, or uses the 
      crossover operator on 2 schedules. When performing mutation or crossover, the 
      schedules are selected with a probability proportional to their fitness, allowing 
      the schedules with high quality to be selected more frequently and make more 
      ``offspring".

      The free parameters in our GA are the size of the population, number of generations 
      to simulate, and proportions of new generations to be made by the copy, random, 
      mutation, and crossover methods.

    \subsubsection{Representation}
      We use the naive representation of schedules. A schedule contains lists of size $|C|$
      containing the room in which each course's exam is to occur and the timing
      (corresponding to a pair or integers representing the day and time) of each course's
      exam. We will see that this representation is far from ideal, but there is no clear 
      alternative.

@@

      The neighbouring schedules are those where only one course meets in either a different
      room, on a different day, or at a different time. 
%      This leads to a total of $ |C|(d+t+|R|-3) $ schedules in the neighbourhood of any given schedule.
      It can be seen that this method is completely deterministic given a starting schedule,
      and will always return a local maximum.

    \subsubsection{Mutate Search}
      Mutate Search starts by generating a random schedule. It makes an alternative schedule
      (by using the same mutate operator made for the genetic algorithm) with 1\% of the
      data randomized and keeps the better schedule. This operation is
      repeated until it has performed as many evaluations as it is allowed to.

@@

    \subsubsection{Parallel Computations}
      Of the 4 methods, only mutate search is not trivially parallelizable. A plausible 
      parallel version of mutate search would generate as many mutated schedules at 
      every step as processors are available and select the fittest result for the next 
      set of mutations. We expect that this would increase the rate at which the best 
      found fitness increases.

  \subsection{Scheduling Problem Instances}
    We generated 2 sets of 3 scheduling problem instances with which to test the 
    above algorithms. The first set of instances were generated with the parameters 
    in the first three columns of the table below. The rooms, courses and students 
    were given arbitrary names and the courses that each student attended were 
    selected with uniform probability from the complete set of courses.

    The second set was generated with a correlation between the courses that each 
    student took, rendering the data-set more true to the sort of data that would 
    be found in real-world situations. The set of courses that each student takes 
    was selected so that around 70\% of their courses were in their (randomly selected) 
    major. This simulates the way in which we expect student bodies to choose their 
    courses, though the true data is likely to exhibit far more structure, due to 
    program requirements, notorious professors, etc.

    The sized of our 3 instances' vary from manageable to large. The total number 
    of different possible schedules in any problem instance is $|C|^{|R|dt}$ which 
    is far too large to conduct a brute force search. The smallest could potentially 
    be solved exactly, with methods not under consideration here, while the others 
    have search spaces much too large for any hope of finding a perfect solution. 
    For the largest, which simulates a UofT exam schedule, a single evaluation of 
    the fitness function takes 0.2 seconds on on a typical desktop computer. This 
    makes it difficult to thoroughly investigate this instance.

    The large data set was created with parameters corresponding to the fall 2009 
	University of Toronto exam schedule, which is the size we expect for a large 
	real-world problem instance.

    \definecolor{light-gray}{gray}{0.87}
    \rowcolors{1}{light-gray}{white}
    \begin{table}[h]
    \begin{small}
    \begin{tabular}{*{7}{c}}
    \hline 
 	  	& S 	& M 	& L 	& $S_{Cor}$ & $M_{Cor}$ & $L_{Cor}$ \\ 
    \hline \hline
     d 		& 7 	& 10	& 7 	& 7 	& 10 	& 7 	\\ 
     t 		& 5		& 8 	& 8 	& 5 	& 8 	& 8 	\\ 
     $|C|$	& 20 	& 200 	& 603 	& 20 	& 200 	& 600 	\\ 
     $|R|$	& 4 	& 10 	& 43 	& 4 	& 10 	& 43	\\ 
     $|S|$	& 50 	& 300 	& 21945 & 50 	& 300 	& 21945	\\ 
     $|L_s|$	& 1-5 	& 1-6 	& 1-5 	& 2-6 	& 2-6 	& 2-6 	\\ 
     \begin{tiny}$\#$Maj\end{tiny}	& N/A 	& N/A 	& N/A 	& 4 	& 10 	& 20 	\\ 
     \textsc{SSS}	& $10^{182}$ & $10^{1840}$ & $10^{6695}$ & $10^{182}$ & $10^{1840}$ & $10^{6689}$ \\
    \hline  
    \end{tabular} 
	\end{small}
	\caption{The six data sets used in our project. $\#$Maj is the number of different 
	         majors used and \textsc{SSS} is an approximation of the search space size.}
	\end{table}

\section{Evaluation}
   \subsection{Selecting Parameters for the GA}
      In order to select parameters for the genetic algorithm (that is, the population, 
      number of generations to simulate, proportions of copies, randoms, mutations, and 
      crossovers for the next generation) we performed tests using the small uncorrelated 
      problem instance.

      \paragraph{Generation Parameters}
        We undertook a search of the space of parameters for forming new generations. 
        Our results for different combinations of parameters can be seen in figure 1. 
        Each point has been averaged over 5 trials. The colour on the plotted surface 
        corresponds to the fitness, where the highest fitnesses are red. The number of 
        copies is fixed at 1, and the proportion of mutations and crossovers are shown 
        on the horizontal axes.

        \begin{figure}[h]
  		  \centering
  		  \includegraphics[width=8cm]{GAParametersData}
  	      \caption{Effectiveness of the GA's proportion parameters}
  	      \label{graph:GAParam1}
		\end{figure}

        The graph indicates that for a fixed number of evaluations, best results are 
        achieved by selecting parameters that have the majority of individuals in the next 
        generation be generated by crossovers of the most fit schedules in the previous 
        generation. In order to achieve the very highest fitness, a few randoms or mutations 
        should also be incorporated.

        This test was done for only the smallest problem instance, and is infeasible for the 
        larger instances. We chose generation proportion parameters to approximate the optimal 
        values seen here, under the assumption that the larger problems exhibited this sort of 
        preference.

  \subsection{Comparison Between the Different Searches}
    We tested the 4 search methods on all 6 of our problem instances. The results 
    for two of the problems, the small and large problems, are shown in figure 2 
    and 3. The medium problem exhibited characteristics very similar to the large problem.

    All methods produced nearly indistinguishable results for the uncorrelated 
    versions. This suggests that it may be possible to use the expected internal 
    correlations more effectively that the methods we have proposed do.

    On the graphs for the small instance, the results shown are averages over many trial 
    runs. Individually, however, the different methods all produced similar results, in 
    particular the ranking of the algorithms was very consistent for any particular problem 
    instance.

    \begin{figure}[h]
  	  \centering
  	  \includegraphics[width=8cm]{SMALL-TYPE2_FullAverageStats}
      \caption{Comparing the four searches on a small data set}
	\end{figure}

	\begin{figure}[h]
  	  \centering
  	  \includegraphics[width=8cm]{HUGE-TYPE2_FullAverageStats}
      \caption{Comparing the four searches on a large data set}
	\end{figure}

	Random search provides a benchmark by which to measure the other 3 techniques. 
	As desired, they all usually perform significantly better.

	Hillclimbing works very well for the small instance, but miserably on the large 
	problem. This is because of the vastly increased number of neighbouring schedules 
	in the large problem. Hillclimbing is able to examine a much smaller portion of 
	search space and wastes its evaluations on too many points that are too similar 
	to each other. A variant that tries some fixed number of schedules in the 
	neighbourhood may perform better.

	The mutate search algorithm performed remarkably well in all tests, finding a 
	schedule at least somewhat better than every other method for every size problem 
	instance.

	The GA is a bit of a disappointment. It works much better than random, meaning 
	that the naive crossover method used does have some value, and it scales well 
	for different sized problem instances, which is an important characteristic. 
	For the large instance, the attained fitness of around 0.4 would be the fitness 
	of a schedule for which every student and room assign 1.4 penalty points to the 
	schedule. This is reasonable, and would perhaps be very exciting if mutate search 
	wasn't so much better than GA for this problem.

\section{Conclusion}
  Any monotone function can be applied to the fitness values to obtain identical 
  results for the 3 non-GA methods, which only need to choose the better fitness 
  and do not use the fitness values in any important way. The GA, however, uses 
  the fitness values to decide the probability of schedules being selected and 
  modified for subsequent generations. It might be possible to reduce or increase 
  the probability of poor functions being selected for mutation or crossover.
  \citeauthor*{pdf:solvingexam} \citeyearpar{pdf:solvingexam} for example, got 
  favourable results by squaring the fitness.

  Modifying the fitness function in a variety of ways (perhaps calculating the 
  total penalty that a schedule gets instead averaging over the values given by 
  each student and room) might improve the  quality of the GA approach to this 
  problem. The fitness function used here could  be modified in many ways, and 
  investigating this would be very interesting.

  Unfortunately, without further investigation and modification, the GA given 
  should not be used. Perhaps a novel form for representing schedules could make 
  a GA approach for this problem perform better than mutate search. We recommend 
  Mutate search for solving problems of this type. A typical PC can generate a 
  relatively good schedule for a large exam scheduling problem overnight using the
  mutate search method. Mutate search is very easy to implement, making it even 
  more attractive.

  Mutate search can be modified to change the rate of mutation and number of attempts 
  at each step, which suggests that much better methods can be developed based on the 
  it. This could be an interesting direction for further research.

\nocite{*}
\bibliographystyle{aaai}
\bibliography{paperBib}
\end{document}
